---
title: "2023 Yukon Egg Thiamine Figures"
author: "Drew Porter"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    fig-width: 8
    fig-height: 6
execute:
  warning: false
  message: false
  echo: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false

# Load packages
library(tidyverse)
library(here)
library(knitr)

# Set global chunk options
knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 6,
  fig.align = 'center',
  dpi = 300,
  warning = FALSE,
  message = FALSE
)

```

```{r}
#| label: import-data

# Read in data
raw_egg_thiamine <- read_rds(file = here("data", "yukon_egg_thiamine_2023.rds"))

# Remove cases with missing nmol_T_g or nmol_T_egg values
egg_thiamine <- raw_egg_thiamine |> 
  filter(!is.na(nmol_T_g) & !is.na(nmol_T_egg))

# Remove cases with missing dry weight corrected values
egg_thiamine_dw <- egg_thiamine |> 
  filter(!is.na(nmol_T_g_dw_corrected))

# Create spawning subsets
egg_thiamine_spawning <- egg_thiamine |> 
  filter(group == "Spawning")

egg_thiamine_dw_spawning <- egg_thiamine_dw |> 
  filter(group == "Spawning")

```

```{r}
#| label: define-global-variables

# Set ggplot theme
theme_set(theme_classic())

# Define global ggplot variables
point_size = 4
jitter_width = 0.2
```

## Egg thiamine concentration by site

```{r}
#| label: concentration-thiamine-site

thiamine_concentration_site <- egg_thiamine |> 
  ggplot(aes(x = site, y = nmol_T_g)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none") +
  facet_grid(~group, scales = "free_x", space = "free")

thiamine_concentration_site
```


## Egg thiamine concentration by group
```{r}
#| label: concentration-thiamine-group

thiamine_concentration_group <- egg_thiamine |> 
  ggplot(aes(x = group, y = nmol_T_g)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none")

thiamine_concentration_group
```

## Total egg thiamine by site
```{r}
#| label: total-thiamine-site

total_thiamine_site <- egg_thiamine |> 
  ggplot(aes(x = site, y = nmol_T_egg)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none") +
  facet_grid(~group, scales = "free_x", space = "free")

total_thiamine_site
```

## Total egg thiamine by group
```{r}
#| label: total-thiamine-group

total_thiamine_group <- egg_thiamine |> 
  ggplot(aes(x = group, y = nmol_T_egg)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none")

total_thiamine_group
```

## Egg size by site
```{r}
#| label: egg-size-site

egg_size_site <- egg_thiamine |> 
  ggplot(aes(x = site, y = g_egg)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none") +
  facet_grid(~group, scales = "free_x", space = "free")

egg_size_site
```

## Egg size by group
```{r}
#| label: egg-size-group

egg_size_group <- egg_thiamine |> 
  ggplot(aes(x = group, y = g_egg)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none")

egg_size_group
```

## Egg percent moisture by site
```{r}
#| label: egg-moisture-site

egg_moisture_site <- egg_thiamine_dw |> 
  ggplot(aes(x = site, y = pct_moisture)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none") +
  facet_grid(~group, scales = "free_x", space = "free")

egg_moisture_site
```

## Egg percent moisture by group
```{r}
#| label: egg-moisture-group

egg_moisture_group <- egg_thiamine_dw |> 
  ggplot(aes(x = group, y = pct_moisture)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none")

egg_moisture_group
```

## Egg percent lipid by site
```{r}
#| label: egg-lipid-site

egg_lipid_site <- egg_thiamine_dw |> 
  ggplot(aes(x = site, y = pct_lipid_wet)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none") +
  facet_grid(~group, scales = "free_x", space = "free")

egg_lipid_site
```

## Egg percent lipid by group
```{r}
#| label: egg-lipid-group

egg_lipid_group <- egg_thiamine_dw |> 
  ggplot(aes(x = group, y = pct_lipid_wet)) +
  geom_boxplot(aes(fill = group),
               outlier.shape = NA) +
  geom_jitter(width = jitter_width) +
  guides(fill = "none")

egg_lipid_group
```


```{r}
#| label: regression

test_fig <- egg_thiamine |> 
  ggplot(aes(x = g_egg, y = nmol_T_g)) +
  geom_point(aes(x = g_egg, y = nmol_T_g, color = group), size = 6) +
  geom_smooth(method = "lm")

test_fit <- lm(nmol_T_g ~ g_egg, data = egg_thiamine)
summary(test_fit)

test_fig

test_fig2 <- egg_thiamine |> 
  ggplot() +
  geom_point(aes(x = g_egg, y = nmol_T_egg, color = group, shape = site), size = 4)

test_fig2


```


``` {r}
#| label: placeholder

spawning <- egg_thiamine |> 
  filter(group == "Spawning")


fig1 <- spawning |> 
  ggplot(aes(x = nmol_T_egg, y = nmol_T_g)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), 
                     limits = c(0, max(spawning$nmol_T_egg + 0.2))) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, max(spawning$nmol_T_g + 5)))

fig1

spawning_fit <- lm(nmol_T_g ~ nmol_T_egg, data = spawning)
summary(spawning_fit)

```

## Dry weight corrected spawner biplot
```{r}
#| label: dw-spawner-fit

dw_spawners <- egg_thiamine_dw |> 
  filter(group == "Spawning")

dw_spawners_plot <- dw_spawners |> 
  ggplot(aes(x = nmol_T_egg, y = nmol_T_g_dw_corrected)) +
  geom_point(aes(shape = site, color = site), size = 3) +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), 
                     limits = c(0, max(dw_spawners$nmol_T_egg + 0.2))) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, max(dw_spawners$nmol_T_g_dw_corrected + 5))) 

dw_spawners_plot

dw_spawners_fit <- lm(nmol_T_g_dw_corrected ~ nmol_T_egg, data = dw_spawners)
summary(dw_spawners_fit)

```

## Percent moisture vs egg size
```{r}
#| label: moisture-vs-size

moisture_size_plot <- egg_thiamine_dw |> 
  ggplot(aes(x = g_egg, y = pct_moisture)) +
  geom_point(size = 2)

moisture_size_plot

```


## nmol T/g (dry) vs nmol T/g (wet) (linear regression, overall)
``` {r}
#| label: dry-vs-wet-thiamine-conc

drywet_fit <- lm(nmol_T_g_dw_corrected ~ nmol_T_g, data = egg_thiamine_dw)
summary(drywet_fit)

drywet_biplot <- ggplot(data = egg_thiamine_dw, aes(x = nmol_T_g, y = nmol_T_g_dw_corrected)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,21)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,50))

drywet_biplot

```


## nmol T/g (dry) vs nmol T/g (wet) (linear regression, spawners)
``` {r}
#| label: dry-vs-wet-thiamine-conc-spawn

drywet_spawn_fit <- lm(nmol_T_g_dw_corrected ~ nmol_T_g, 
                       data = egg_thiamine_dw_spawning)
summary(drywet_spawn_fit)

drywet_spawn_biplot <- ggplot(data = egg_thiamine_dw_spawning,
                              aes(x = nmol_T_g, y = nmol_T_g_dw_corrected)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,10)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,20))

drywet_spawn_biplot

```

## nmol T/g (dry) vs site (box plot)
``` {r}
#| label: dry-conc-site

dryconc_site <- ggplot(data = egg_thiamine_dw, aes(x = site, y = nmol_T_g_dw_corrected)) +
  geom_boxplot(aes(fill = group)) +
  geom_jitter(width = jitter_width)

dryconc_site

```

## nmol T/g (dry) vs group (box plot)
``` {r}
#| label: dry-conc-group

dryconc_group <- ggplot(data = egg_thiamine_dw, aes(x = group, y = nmol_T_g_dw_corrected)) +
  geom_boxplot(aes(fill = group)) +
  geom_jitter(width = jitter_width)

dryconc_group

```

## nmol T/g (dry) vs nmol T/individual egg (linear regression, overall)
``` {r}
#| label: dry-vs-egg-thiamine-conc

dryegg_fit <- lm(nmol_T_g_dw_corrected ~ nmol_T_egg, data = egg_thiamine_dw)
summary(dryegg_fit)

dryegg_biplot <- ggplot(data = egg_thiamine_dw, aes(x = nmol_T_egg, y = nmol_T_g_dw_corrected)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,50))

dryegg_biplot

```


## nmol T/g (dry) vs nmol T/individual egg (linear regression, spawners)
``` {r}
#| label: dry-vs-egg-thiamine-conc-spawn

dryegg_spawn_fit <- lm(nmol_T_g_dw_corrected ~ nmol_T_egg, 
                       data = egg_thiamine_dw_spawning)
summary(dryegg_spawn_fit)

dryegg_spawn_biplot <- ggplot(data = egg_thiamine_dw_spawning,
                              aes(x = nmol_T_egg, y = nmol_T_g_dw_corrected)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,20))

dryegg_spawn_biplot

```


## nmol T/g (wet) vs nmol T/individual egg (linear regression, overall)
``` {r}
#| label: wet-vs-egg-thiamine-conc

wetegg_fit <- lm(nmol_T_g ~ nmol_T_egg, data = egg_thiamine)
summary(wetegg_fit)

wetegg_biplot <- ggplot(data = egg_thiamine, aes(x = nmol_T_egg, y = nmol_T_g)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,25))

wetegg_biplot

```

## nmol T/g (wet) vs nmol T/individual egg (linear regression, spawners)
``` {r}
#| label: wet-vs-egg-thiamine-conc-spawn

wetegg_spawn_fit <- lm(nmol_T_g ~ nmol_T_egg, data = egg_thiamine_spawning)
summary(wetegg_spawn_fit)

wetegg_spawn_biplot <- ggplot(data = egg_thiamine_spawning,
                              aes(x = nmol_T_egg, y = nmol_T_g)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,10))

wetegg_spawn_biplot

```

```{r}
#| label: save-plots
#| include: false

plots <- list(
  thiamine_concentration_site = thiamine_concentration_site,
  thiamine_concentration_group = thiamine_concentration_group,
  total_thiamine_site = total_thiamine_site,
  total_thiamine_group = total_thiamine_group,
  egg_size_site = egg_size_site,
  egg_size_group = egg_size_group,
  egg_moisture_site = egg_moisture_site,
  egg_moisture_group = egg_moisture_group,
  egg_lipid_site = egg_lipid_site,
  egg_lipid_group = egg_lipid_group
)


for (plot_name in names(plots)) {
  ggsave(paste0(plot_name, ".png"), 
       plot = plots[[plot_name]],
       path = here("output"),
       height = 6,
       width = 8,
       units = "in",
       dpi = 300)
}
```
